---
title: "Homework 2"
author: "Irene Yang"
date: "2/21/2018"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
## Setting Up

<span style="color:red">Installed packages: `car` and `quantreg`.  Note: Responded "n" to prompt "Do you want to install from sources the package which needs compilation?"</span>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(quantreg)
library(dplyr)
```

## Homework on the `Davis` dataset in the `car` package - using the `dplyr` package as much as possible.

### **Question 1:** What kind of R object is the `Davis` dataset?

```{r H2Q1}
tbl_df(Davis)
```

<span style="color:red">The `Davis` dataset is a dataframe made up of 200 rows and five columns. The columns describe the sex, weight, height, repwt (reported weight), repht (reported height) of men and women who exercise regularly. Sex is a factor vector and the remaining columns are integer vectors.</span>  

### **Question 2:** How many observations are in the `Davis` dataset?

```{r H2Q2}
count(Davis)
```
<span style="color:red">There are 200 observations in this dataset.</span.

### **Question 3:** For reported weight, how many observations have a missing value?

```{r H2Q3}
summary(Davis)
```

<span style="color:red">There are 17 missing values in the repwt column.</span>

### **Question 4:** How many observations have no missing values? _(HINT: find complete cases)_

```{r H2Q4}
Davis %>% count(complete.cases(Davis))
```

<span style="color:red">181 observations have no missing values.</span>

### **Question 5:** How many females are in this subset?

```{r Create subset containing only females}
count(Davis, sex == "F")
```

<span style="color:red">There are 112 females in this subset.</span>

### **Question 6:** What is the average BMI for these individuals?

```{r Compute BMI variable}
Davis_BMI <- mutate(Davis, height_m = height/100, BMI = weight/height_m^2)
summary(Davis_BMI)
```

<span style="color:red">The average BMI is 24.70.</span>

### **Question 7:** How do these individuals fall into the BMI categories (what are the frequencies and relative %'s)?

```{r BMI categories}
Davis_BMICategory <- Davis_BMI %>%
  mutate(BMICategory = case_when(BMI < 18.5 ~ "Underweight",
                                 BMI > 18.5 & BMI < 25 ~ "Normal",
                                 BMI > 25 & BMI < 30 ~ "Overweight",
                                 BMI >= 30 ~ "Obese",
                                 BMI = NA ~ "Missing"))
Davis_BMICategory %>%
  group_by(BMICategory) %>%
  summarize(n = n()) %>%
  mutate(relfreq = n / sum(n) * 100)
```

<span style="color:red">Frequencies and relative frequencies are listed above.</span>
  
### **Question 8:** Create a histogram of BMI. 

```{r Histogram of BMI}
# log10 transformation performed in an attempt to normalize the values (extremely high BMI of 510.93!)
library(ggplot2)
Davis_BMICategory %>%
  ggplot() +
  geom_histogram(aes(log10(BMI)), binwidth = .05, color = "black", fill = "grey") +
  ggtitle("BMI in Davis Dataset")
```

<span style="color:red">Even with the log transformation, there is an extreme outlier.</span>

### **Question 9:** Create side-by-side boxplots of the BMI distributions by gender

```{r Side-by-side boxplots BMI by gender}
# First, extreme outlier (BMI of 510.93) removed.
Davis_BMI_new <- filter(Davis_BMICategory, BMI < 510) %>%
  select(sex, BMI, BMICategory)
  ggplot(Davis_BMI_new, aes(x = sex, y = BMI)) +
  geom_boxplot()+
  ggtitle("BMI by Gender")
```

### **Question 10:** Create a clustered bar chart of the BMI categories by gender

```{r Clustered bar chart BMI category by gender}
#Use data frame that has been stripped of the extreme outlier.
Davis_BMI_new %>%
  select(sex, BMICategory)
  ggplot(Davis_BMI_new, aes(sex)) +
           geom_bar(aes(fill = BMICategory), position = "dodge")
```

